Tropical Andes montane and lowland forest bird and mammal species GBIF

### 2024-05-16 code/L0/L0_TropicalAndes_frugivores_GBIF.Rmd
update file paths, add date data was downloaded, add statement not to re-request data if file exists, update summary with custom function, add package citation and session info at end

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

@@

# Read in Frugivoria species list
```{r}
frugivore_species <- read.csv(file.path(data_path,"TropicalAndes_Frugivoria_frugivore_traits_species.csv"))
```

# Match species names to GBIF
```{r}
gbif_taxon_keys <- frugivore_species$x %>% 
name_backbone_checklist() %>% # match to backbone 
filter(!matchType == "NONE") %>% # get matched names
pull(usageKey) 
```

# Download data
```{r}
occ_download(
pred_in("taxonKey", gbif_taxon_keys),
pred("hasCoordinate", TRUE),
pred("hasGeospatialIssue", FALSE),
pred_in("country",c("EC","CO","VE", "PE", "BO")), 
format = "SIMPLE_CSV"
)
```

# Retrive GBIF download, save file, and load into r
```{r}
d <- occ_download_get("0009771-231002084531237", path = output_path, overwrite = TRUE) %>%
  occ_download_import()
```

# Subset occurrence to Tropical Andes IUCN Montane and Lowland forest

## Read in Forest shape
```{r}
TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path, "Forest_sf.shp"), layer = "Forest_sf")
```

@@

@@

```{r}
print("The number of records is")
nrow(TropicalAndes_frugivore_occ_forest)
print("The number of species is")
length(unique(TropicalAndes_frugivore_occ_forest$species))
print("The number of genera is" )
length(unique(TropicalAndes_frugivore_occ_forest$genus))
print("The number of families is")
length(unique(TropicalAndes_frugivore_occ_forest$family))
```

@@

# Write data to csv
```{r}
write.csv(TropicalAndes_frugivore_occ_forest, file = file.path(output_path,"TropicalAndes_GBIF_frugivore_occ.csv"))

write.csv(GBIF_occ_SpeciesList, file = file.path(output_path,"TropicalAndes_GBIF_frugivore_occ_species.csv"))
```

>>>>

# Set file paths
```{r}
data_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

@@

# Read in Frugivoria species list
```{r}
frugivore_species <- read.csv(file.path(data_path_L0,"TropicalAndes_Frugivoria_frugivore_traits_species.csv"))
```

# Download data
Data downloaded on 2023-08-02
```{r}
file_name <- "0009771-231002084531237.zip"

if (!file.exists(file.path(output_path_L0, file_name))) {
    # Match species names to GBIF  
    gbif_taxon_keys <- frugivore_species$x %>% 
      name_backbone_checklist() %>% # match to backbone 
      filter(!matchType == "NONE") %>% # get matched names
      pull(usageKey) 
    # Make GBIF request
    occ_download(
      pred_in("taxonKey", gbif_taxon_keys),
      pred("hasCoordinate", TRUE),
      pred("hasGeospatialIssue", FALSE),
      pred_in("country",c("EC","CO","VE", "PE", "BO")), 
      format = "SIMPLE_CSV"
    )
    # Download GBIF data
  d <- occ_download_get("0009771-231002084531237",
                        path = output_path_L0, overwrite = TRUE) %>%
    occ_download_import()
  } else {
    cat("File already exists:", file_name, "\n")
    d <- occ_download_get("0009771-231002084531237",
                        path = output_path_L0, overwrite = TRUE) %>%
                        curveocc_download_import()
  }
```

# Subset occurrence to Tropical Andes IUCN Montane and Lowland forest
## Read in Forest shape
```{r}
TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path, "Forest_sf.shp"), layer = "Forest_sf")
```

@@

# Fix column type 
```{r}
TropicalAndes_frugivore_occ_forest$gbifID <- as.character(TropicalAndes_frugivore_occ_forest$gbifID)
TropicalAndes_frugivore_occ_forest$geometry <- NULL
```

@@

```{r}
source("~/GitHub/neotropical_plants/code/data_summary.R")

data_summary(TropicalAndes_frugivore_occ_forest, TropicalAndes_frugivore_occ_forest$species, TropicalAndes_frugivore_occ_forest$genus, TropicalAndes_frugivore_occ_forest$family)
```

@@

# Write data to csv
```{r}
write.csv(TropicalAndes_frugivore_occ_forest, file = file.path(output_path_L0,"TropicalAndes_GBIF_frugivore_occ.csv"))

write.csv(GBIF_occ_SpeciesList, file = file.path(output_path_L0,"TropicalAndes_GBIF_frugivore_occ_species.csv"))
```

# Package Citations and Session Info
```{r}
library(report)
report::cite_packages()

devtools::session_info()
```
